Leadership across cultures and contexts: an exploratory analysis

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Added a senetance in discussion.

We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful groups. Variables were clustered together if they were positively correlated. Leadership and high social status are expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019]. Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies, prestige or dominance; for example, @van_vugt_dual_2019 [p. 1] "argue that that there are two leadership styles: prestige style and dominance style" and @cheng_two_2013 [p. 119] in focusing on dominance-prestige as two distinct leadership strategies suggest that, "that human social stratification is characterized by the co-occurrence of both strategies."

Exploratory cluster analyses reported here suggest multiple features of leader qualities and leader functions across cultures and contexts. Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, mediate problems, and provide prosocial services to the group, all while conforming to social and cultural pressures.

>>>>

We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful groups. Variables were clustered together if they were positively correlated. Leadership and high social status are expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019]. Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: prestige or dominance. @van_vugt_dual_2019 [p. 1], for example, "argue that that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p. 119] in focusing on dominance vs. prestige as two distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies."

We acknowledge that the dominance-prestige theory as formulated by @henrich_evolution_2001 has significantly advanced evoluytioanry approaches to leadership, however, exploratory cluster analyses reported here suggest multiple features of leader qualities and leader functions across cultures and contexts. Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, mediate problems, and provide prosocial services to the group, all while conforming to social and cultural pressures.
